market-justice-preferences

### 2024-06-24 01-introduction.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities. Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system  [@bourdieu_reproduction_1990]. A large part of the research in this field at an international  level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the socio-economic and cultural gaps of origin. Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities even perceived at the school context? Are they rejected by the students, particularly those who are worst-off in socioeconomic terms? Or, Is there evidence at the school level that social inequalities are tolerated and even justified? [@batruch_belief_2022; @wiederkehr_belief_2015].

@@

In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system. In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].\
Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022]. Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.

Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022]. We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes. To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.

The Chilean case is particularly intriguing for studying market justice preferences. This country is characterized by acute and persistent economic inequality, which stands out in Latin America and among OECD countries [@flores_top_2020]. In Chile, the poorest 50% captures only 10% of the total income and has negative wealth, while the richest 1% receives almost 27% of the income and holds 49.6% of the wealth [@chancel_world_2022]. Much of this inequality has been attributed to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors. Such reforms where introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018]. This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020]. In health, although the majority of the population uses the public insurance system (78.9%), the pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024). In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).

>>>>

Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities. Despite this, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system  [@bourdieu_reproduction_1990]. A large part of the research in this field at an international  level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the socio-economic and cultural gaps of origin. Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities even perceived at the school context? Are they rejected by the students, particularly those who are worst-off in socioeconomic terms? Or, Is there evidence at the school level that social inequalities are tolerated and even justified? [@batruch_belief_2022; @wiederkehr_belief_2015].

@@

In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system. In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023]. Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022]. Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.

Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger market justice preferences, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022]. We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes. To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.

The Chilean case is particularly intriguing for studying market justice preferences. This country is characterized by acute and persistent economic inequality, which stands out in Latin America and among OECD countries [@flores_top_2020]. In Chile, the poorest 50% captures only 10% of the total income and has negative wealth, while the richest 1% receives almost 27% of the income and holds 49.6% of the wealth [@chancel_world_2022]. Much of this inequality has been attributed to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors. Such reforms where introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018]. This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020]. In health, although the majority of the population uses the public insurance system (78.9%), 15.3% are served by private insurers (Observatorio Social, 2023). The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024). In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Ministerio de Educación, 2023).

### 2024-06-24 04-analisis.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4,
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               easystats)

@@

This contrasts with the pattern observed for these items in the Chilean adult population (see @fig-elsoc). While both age groups tend to disagree that higher-income individuals should access better health, pension, and education services, this tendency is more pronounced in adults than in students, who show greater agreement with market justice preferences in these areas. Additionally, the hierarchy of these domains differs between groups; adults justify market justice least in education, followed by health, and then pensions, whereas students justify market principles less in health, then pensions, and lastly in education.

@@

data_rec %>% 
  select(just_salud, just_pension, just_educ) %>% 
  sjPlot::plot_stackfrq(geom.size = 0.5,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Health", "Pensions", "Education")) +
 # theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
                             " (N = ",dim(data_rec)[1],")"))

@@

plot + 
  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
                             " (N = ",dim(data_rec)[1],")"))
```

@@

ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
  labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
                             " (N = ",dim(data_rec)[1],")"))
```

@@

texreg::htmlreg(l = list(modelos_health[[2]],
                         modelos_health[[3]],
                         modelos_pension[[2]],
                         modelos_pension[[3]],
                         modelos_educ[[2]],
                         modelos_educ[[3]]),   
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
                custom.model.names = rep(c("Model 1", "Model 2"), 3),
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= Non-university)" = 6:7,
                                "Socioeconomic level (Ref.= Low)" = 9:10,
                                "Achievement score (Ref.= Low)" = 11:12),
                custom.gof.rows = list("Controls"=rep("Yes",6)),
                include.loglik = FALSE, include.aic = FALSE)

@@

texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                custom.model.names = c("Model 1",
                                       "Model 2",
                                       "Model 3",
                                       "Model 4"),
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= Non-university)" = 7:8,
                              "Socioeconomic level (Ref.= Low)" = 10:11,
                              "Achievement score (Ref.= Low)" = 12:13),
                custom.gof.rows = list("Controls"=rep("Yes",4),
                                       "Deviance" = c(deviance1,
                                                    deviance2,
                                                    deviance4,
                                                    deviance5),
                                       "Deviance Test (p)" = c(test.pvalues1,
                                                             test.pvalues2,
                                                             test.pvalues4,
                                                             test.pvalues5)),
                include.loglik = FALSE,include.aic = FALSE)

@@

collapse_rows[c(-4),] %>% 
  as_tibble() %>% 
  kbl(.,
      align = "lccccc", booktabs = T,
      col.names =  c("Variable", 
                     "School talent", 
                     "School effort",
                     "Social talent",
                     "Social effort", 
                     "Deservingness")) %>% 
  add_header_above(c(" " = 1,
                     "Market Justice Preferences"=5), 
                   bold = T) %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>% 
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05", 
                           notation = "none")

@@

plot_interact<- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
  ylim(1,4)+
  labs(x = "Deservingness", y="Market Justice Preferences", colour = "School Simce")+
  theme_bw()
```

@@

plot_interact

>>>>

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4,
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               easystats,
               ggeffects)

@@

@@

data_rec %>% 
  select(just_salud, just_pension, just_educ) %>% 
  sjPlot::plot_stackfrq(geom.size = 0.5,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Health", "Pensions", "Education")) +
 # theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")

@@

plot + 
  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
       labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
```

@@

ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
  labs(caption = "Source: own elaboration based on First Study of Civic Education 2017")
```

@@

texreg::htmlreg(l = list(modelos_health[[2]],
                         modelos_health[[3]],
                         modelos_pension[[2]],
                         modelos_pension[[3]],
                         modelos_educ[[2]],
                         modelos_educ[[3]]),   
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars. Source: own elaboration based on First Study of Civic Education 2017.",
                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
                custom.model.names = rep(c("Model 1", "Model 2"), 3),
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= Non-university)" = 6:7,
                                "Socioeconomic level (Ref.= Low)" = 9:10,
                                "Achievement score (Ref.= Low)" = 11:12),
                custom.gof.rows = list("Controls"=rep("Yes",6)),
                include.loglik = FALSE, include.aic = FALSE)

@@

texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                custom.model.names = c("Model 1",
                                       "Model 2",
                                       "Model 3",
                                       "Model 4"),
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars. Source: own elaboration based on First Study of Civic Education 2017.",
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= Non-university)" = 7:8,
                              "Socioeconomic level (Ref.= Low)" = 10:11,
                              "Achievement score (Ref.= Low)" = 12:13),
                custom.gof.rows = list("Controls"=rep("Yes",4),
                                       "Deviance" = c(deviance1,
                                                    deviance2,
                                                    deviance4,
                                                    deviance5),
                                       "Deviance Test (p)" = c(test.pvalues1,
                                                             test.pvalues2,
                                                             test.pvalues4,
                                                             test.pvalues5)),
                include.loglik = FALSE,include.aic = FALSE)

@@

collapse_rows[c(-4),] %>% 
  as_tibble() %>% 
  kbl(.,
      align = "lccccc", booktabs = T,
      col.names =  c("Variable", 
                     "School talent", 
                     "School effort",
                     "Social talent",
                     "Social effort", 
                     "Deservingness")) %>% 
  add_header_above(c(" " = 1,
                     "Market Justice Preferences"=5), 
                   bold = T) %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>% 
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05. Source: own elaboration based on First Study of Civic Education 2017.", 
                           notation = "none")

@@

plot_interact <- ggpredict(model = reg26.2_bienestar, terms = c("merito_soc", "simce")) %>% 
  filter(group != "Medium") %>% 
  plot() +
  ylim(1,4) +
  labs(title = NULL,
       x = "Deservingness", 
       y ="Market Justice Preferences", 
       colour = "School Simce",
       caption = "Source: own elaboration based on First Study of Civic Education 2017")
```

@@

plot_interact +
  theme_bw()

### 2024-06-24 05-discussion.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

The previous analyses were aimed at delving into the relationship between perceptions of meritocracy and preferences for market justice among students. Several models estimated how students' beliefs about the rewards for talent and effort in both society and school environments influence their attitudes towards access to social services like health, education, and pensions. The models also examined the impact of students' socioeconomic status and school achievement on these preferences, analyzing whether perceptions of meritocracy moderate these effects. The findings provide insights into the complex interplay between meritocratic beliefs, socioeconomic factors, and attitudes towards social justice.

@@

Hypothesis $H_x$ [Falta hipótesis school achievement] held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results. Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice. This finding is one of the most strongest and consistent throughout the models and it is relevant given the centrality of school performance in educational institutions. One possible explanation of this link is that students in high-achieving schools may be more exposed to discussions and to a deeper understanding about inequality and social justice, leading to a critical perspective towards the weight of market mechanism in the access to social services.

Hypotheses $H_6$ and $H_7$ proposed that family status would moderate the effect of meritocratic perceptions on market justice preferences. Hypothesis $H_6$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the society level. This means that the positive impact of the perception of social meritocracy on market justice preferences have a lower effect in students with parents with higher education and also in those with more books at home. Something similar occurs with hypothesis $H_7$, where now school level status moderates the effect of meritocracy variables, which in multilevel framework si referred as to cross-level interaction. The negative interaction in this context means that in schools of low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger. These relations indicate that, although those who perceive society as meritocratic have greater preferences towards market justice, this influence is lower among those with a higher socioeconomic level. This finding highlights the complex interaction between socioeconomic status, school context, meritocratic beliefs, and preferences toward market justice. The moderating role of meritocratic perceptions suggests that students' views on social justice are shaped by their socioeconomic background and beliefs about how society rewards effort and talent. This underlines the importance of addressing structural inequalities and individual beliefs in educational and social policies.

>>>>

The previous analyses aimed to explore the relationship between perceptions of meritocracy and preferences for market justice among students in Chilean society. Several models estimated how students' beliefs about the rewards for talent and effort in both societal and school contexts influence their attitudes toward access to social services like health, education, and pensions based on pay capacity. Additionally, the models examined the impact of students' socioeconomic status and school achievement on these preferences, assessing whether perceptions of meritocracy moderate these effects. The findings offer insights into the complex interplay between meritocratic beliefs, socioeconomic factors, and attitudes toward market justice criteria.

@@

Hypothesis $H_4$ held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such school could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results. Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice. This finding is one of the most strongest and consistent throughout the models and it is relevant given the centrality of school performance in educational institutions. One possible explanation of this link is that students in high-achieving schools may be more exposed to discussions and to a deeper understanding about inequality and social justice, leading to a critical perspective towards the weight of market mechanism in the access to social services.

Hypotheses $H_5$ and $H_6$ proposed that socioeconomic status would moderate the effect of meritocratic perceptions on market justice preferences. Hypothesis $H_5$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the society level. This means that the positive impact of the perception of social meritocracy on market justice preferences have a lower effect in students with parents with higher education and also in those with more books at home. Something similar occurs with hypothesis $H_6$, where now school level status moderates the effect of meritocracy variables, which in multilevel framework si referred as to cross-level interaction. The negative interaction in this context means that in schools of low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger. These relations indicate that, although those who perceive society as meritocratic have greater preferences towards market justice, this influence is lower among those with a higher socioeconomic level. This finding highlights the complex interaction between socioeconomic status, school context, meritocratic beliefs, and preferences toward market justice. The moderating role of meritocratic perceptions suggests that students' views on social justice are shaped by their socioeconomic background and beliefs about how society rewards effort and talent. This underlines the importance of addressing structural inequalities and individual beliefs in educational and social policies.

$H_7$ Falta solo esto

### 2024-06-24 06-conclusion.qmd
Revision Andreas y Render completo
Revision Andreas:
- Falta discussion de H7
- Referencias que faltan agregarlas a Zotero (Kevin)

The core of this paper was the association between meritocratic perceptions and preferences for market justice. In general, we find that those who perceive that effort and talent/intelligence are rewarded, are more willing to agree with that richer individuals can have better health, education, and pensions. This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities [@mijs_paradox_2019; @darnon_where_2018], but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it. Although there are still several open questions, as for instance differences between meritocratic perceptions in school and in the society at large, this raises up the relevance of what is done (or not) in schools in order to challenge the meritocratic ideals.

The consideration of the school context opened some caveats in this research. We found contrary evidence for the association between status (as measured by average education), school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences. This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms. A possible explanation is that education contributes to form a crirical view of markets in redistribution, nevertheless further reserach in this area is needed in order to assert the specific mechanisms that might be playing a role. All in all, from this study we know that school contexts are not trivial to this regard, calling for larger attention to school age research on the formation of inequality preferences and beliefs.

>>>>

The core of this paper was the association between meritocratic perceptions and preferences for market justice. In general, we find that those who perceive that effort and talent/intelligence are rewarded, are more willing to agree with that richer individuals can have better health, education, and pensions. This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities (Mijs, 2019; Darnon 2017), but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it. Although there are still several open questions, as for instance differences between meritocratic perceptions in school and in the society at large, this raises up the relevance of what is done (or not) in schools in order to challenge the meritocratic ideals.

The consideration of the school context opened some caveats in this research. We found contrary evidence for the association between status, school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences. This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms. A possible explanation is that education contributes to form a crirical view of markets in redistribution, nevertheless further reserach in this area is needed in order to assert the specific mechanisms that might be playing a role. All in all, from this study we know that school contexts are not trivial to this regard, calling for larger attention to school age research on the formation of inequality preferences and beliefs.
